Routines and Regret: An Examination of Behavioral Norms and Emotional Responses

### 2024-02-15 paper/paper.qmd
modify

#### Workspace set-up ####
library(ggplot2)
library(scales)
library(knitr)
library(MBESS)
library(psych)
library(dplyr) 
library(Hmisc)
library(effsize)
library(jmv)
library(Rcpp)
library(reshape2)
library(readr)
library(janitor)
library(tidyverse)
```

@@

we use R [@citeR] for all data wrangling and analysis and R packages ggplot2 [@ggplot2], scales [@scales],knitr [@knitr], MBESS [@MBESS], psych [@psych], dplyr [@dplyr], Hmisc [@Hmisc], effsize [@effsize], jmv [@jmv], Rcpp [@Rcpp], reshape2 [@reshape2], readr [@readr], janitor [@janitor], tidyverse [@tidyverse] to read in, clean and modify the data, as well as view statistical information by conducting binomial test. Furthermore, produce the tables and figures to show our results.

The remainder of this paper is structured as follows. @sec-data In the Data part, we detail the methodology behind our data collection and analysis, and describe the survey data used. @sec-results In the Result part, We employ plots and textual analysis to present our findings in two scenario, illustrating the relationship between routine, regret, social norms, and negative affect across different contexts. @sec-dicussion Moving beyond the raw data, the Discussion part interprets our findings within a broader psychological and societal framework. @sec-conclusion In the conclusion part, we summarize our key discoveries, how our results contribute to the existing body of knowledge on regret and routine behaviors. @sec-appendixIn the appendix, the processes of simulating, reading, cleaning, gathering statistical insights, and testing the data are comprehensively outlined.The figures and tables are based on the data collected and detailed in the appendix section that aggregates statistical information.

@@

There were 342 participants in the replication experiment (127 males, 215 females, Mage = 39.93, SDage = 12.88). 
, which are listed below. 
Given scenario: Mr. Jones almost never takes hitch-hikers in his car. Yesterday he gave a man a ride and was robbed. Mr. Smith frequently takes hitch-hikers in his car. Yesterday he gave a man a ride and was robbed. (Lucas Kutscher & Gilad Feldman (2019))

@@

## Part 1: hitchhiker - scenario
Table 1 is a replication of table 1 in the original paper, and it shows a total of 342 counts and proportions for perceived regret, social norms, and negative affect in the hitch-hiking Experiment 1.

@@

Figure 1 is a replication of figure 1 in the original paper, It presents the proportion of participants’ results clearly with respect to regret, social descriptive norms, social injunctive norms and negative affect. According to the bar plot, there exists a high degree of consistency in participants' responses across four different aspects, with more than 90% of participants agreeing on the same point of view.

@@

## Part 2: Car accident scenario
Figure2 is a replication of figure2 in the original paper, showing the counts and proportions of perceived regret and luck for car accident events.

Table2 shows that most participants regretted the exception (277, 81%) more strongly than the rule (65,19%; p < .001, d = 1.58). The same result can be seen more intuitively in figure2. the result for the exception event represented by Mr. White is much higher for both Regret and Luck than for the Routine event represented by Mr. Adams.

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

@@

## Weaknesses and next steps

@@

## Read raw data

We acquired a CSV file containing raw data relevant to 'Routines and Regret: An Examination of Behavioral Norms and Emotional Responses.' Subsequently, we inspected the first six rows and assessed the demographic details.

```{r}
#| echo: false
#| warning: false
#| message: false

@@

#Check the top six rows of data of interest
dataset_selected <- data %>% select(Sc1_regret, sc1_socnorms1, sc1_socnorms2, sc1_combinednorms) 
kable(head(dataset_selected))

head(data)

# To check exclusions based on pre-registration criteria uncomment the following
# Criteria:
# Serious lower than 5
# English understanding lower than 5
# data <- data[ which(4<data$serious 
#                     & data$engunder > 4), ]

@@

## Clean and test data for statistical information

123

The top 6 rows are shown by @tbl-cleanedata
```{r}
#| echo: false
#| warning: false
#| include: false

@@

@@

@@

###################
# EXPERIMENT 2 (Car Accident-Scenario)

@@

# Descriptives for main variables
jmv::descriptives(
  data=data,
  vars=c(
    "Sc2_regret",
    "Sc2_random_1",
    "Sc2_random_2",
    "Sc2_lucky"),
  freq=TRUE)

# binomial Z
jmv::propTest2(
  data=data,
  vars=c(
    "Sc2_regret",
    "Sc2_lucky"),
  ci=TRUE)
```

@@

>>>>

#### Workspace set-up ####
library(ggplot2)
library(scales)
library(knitr)
library(MBESS)
library(psych)
library(dplyr) 
library(Hmisc)
library(effsize)
library(jmv)
library(Rcpp)
library(reshape2)
library(readr)
library(janitor)
library(tidyverse)
library(kableExtra)
```

@@

For all data wrangling and analysis, we utilize R [@citeR], complemented by a suite of R packages: ggplot2 [@ggplot2] for visualization, dplyr [@dplyr] for data manipulation, along with others such as scales [@scales], knitr [@knitr], kableExtra [@kableExtra] and tidyverse [@tidyverse] for tasks including data cleaning and modification. Additionally, packages like MBESS [@MBESS], psych [@psych], Hmisc [@Hmisc], effsize [@effsize], jmv [@jmv], Rcpp [@Rcpp], reshape2 [@reshape2], readr [@readr], and janitor [@janitor] assist in processing data and viewing statistical information. Importantly, the creation of tables and figures is directly based on the statistical data obtained, ensuring our results are accurately represented and effectively communicated.

The remainder of this paper is structured as follows. @sec-data In the Data part, we detail the methodology behind our data collection and analysis, and describe the survey data used. @sec-results In the Result part, We employ plots and textual analysis to present our findings in two scenario, illustrating the relationship between routine, regret, social norms, and negative affect across different contexts. @sec-discussion Moving beyond the raw data, the Discussion part interprets our findings within a broader psychological and societal framework. @sec-conclusion In the conclusion part, we summarize our key discoveries, how our results contribute to the existing body of knowledge on regret and routine behaviors. @sec-appendix In the appendix, the processes of simulating, reading, cleaning, gathering statistical insights, and testing the data are comprehensively outlined.The figures and tables are based on the data collected and detailed in the appendix section that aggregates statistical information.

@@

There were 342 participants in the replication experiment (127 males, 215 females, Mage = 39.93, SDage = 12.88). 
, which are listed below. 
Given scenario: Mr. Jones almost never takes hitch-hikers in his car. Yesterday he gave a man a ride and was robbed. Mr. Smith frequently takes hitch-hikers in his car. Yesterday he gave a man a ride and was robbed [@kutscher_impact_2019].

@@

## Part 1: hitchhiker - scenario
@tbl-frequency_and_proportion_exp1 is a replication of table 1 in the original paper, and it shows a total of 342 counts and proportions for perceived regret, social norms, and negative affect in the hitch-hiking Experiment 1.

@@

# R's way of doing binomial
# we need to count, while excluding NAs
x1 <- sum((data$Sc1_regret[!is.na(data$Sc1_regret)])=="Exception Jones")
n1 <- length(data$Sc1_regret[!is.na(data$Sc1_regret)])
prop.test(x1, n1, p=0.5, correct = FALSE)
result <- prop.test(x1, n1, p = 0.5, correct = FALSE)
p1 <- result$estimate["p"]

x2 <- sum((data$sc1_socnorms1[!is.na(data$sc1_socnorms1)])=="Exception Jones")
n2 <- length(data$sc1_socnorms1[!is.na(data$sc1_socnorms1)])
prop.test(x2, n2, p=0.5, correct = FALSE)
result <- prop.test(x2, n2, p = 0.5, correct = FALSE)
p2 <- result$estimate["p"]

x3 <- sum((data$sc1_socnorms2[!is.na(data$sc1_socnorms2)])=="Exception Jones")
n3 <- length(data$sc1_socnorms2[!is.na(data$sc1_socnorms2)])
prop.test(x3, n3, p=0.5, correct = FALSE)
result <- prop.test(x3, n3, p = 0.5, correct = FALSE)
p3 <- result$estimate["p"]

x4 <- sum((data$sc1_combinednorms[!is.na(data$sc1_combinednorms)])=="Exception Jones")
n4 <- length(data$sc1_combinednorms[!is.na(data$sc1_combinednorms)])
prop.test(x4, n4, p=0.5, correct = FALSE)
result <- prop.test(x4, n4, p = 0.5, correct = FALSE)
p4 <- result$estimate["p"]

# create a data frame to save the data
table_data <- data.frame(
  ` ` = c("Routine Smith", "Exception Jones"),
  Regret_Counts = c(n1 - x1, x1),
  Regret_Percent = c(round(100*(1 - p1), digits = 1), round(100*p1, digits = 1)),
  Social_Norm_Injunctive_Counts = c(n2 - x2, x2),
  Social_Norm_Injunctive_Percent = c(round(100*(1 - p2), digits = 1), round(100*p2, digits = 1)),
  Social_Norm_Descriptive_Counts = c(n3 - x3, x3),
  Social_Norm_Descriptive_Percent = c(round(100*(1 - p3), digits = 1), round(100*p3, digits = 1)),
  Negative_Affect_Counts = c(n4 - x4, x4),
  Negative_Affect_Percent = c(round(100*(1 - p4), digits = 1), round(100*p4, digits = 1))
)
```

```{r}
#| echo: false
#| label: tbl-frequency_and_proportion_exp1
#| tbl-cap: "Frequency and proportion of regret, social norms, and negative affect for exp1."
# create the table with kable()
kable(table_data, format = "latex", col.names = c(" ", "Regret", "%", "Soc_norm1", "%", "Soc_norm2", "%", "Negetive affect", "%"), align = 'c')
```

@fig-proportion_exp1 is a replication of figure 1 in the original paper, It presents the proportion of participants’ results clearly with respect to regret, social descriptive norms, social injunctive norms and negative affect. According to the bar plot, there exists a high degree of consistency in participants' responses across four different aspects, with more than 90% of participants agreeing on the same point of view.

@@

```{r}
#| echo: false
#| label: fig-proportion_exp1
#| fig-cap: "Proportion of regret, social norms, and negative affect for exp1."
# Data in percentage for Smith and Jones we gained from the table we created
# Ensure that the percentages are divided by 100 to be in the correct format
data2 <- data.frame(
  Category = c('Regret', 'Social norms (injunctive)', 'Social norms (descriptive)', 'Negative affect'),
  Smith = c(7.9 / 100, 4.7 / 100, 90.6 / 100, 7.3 / 100), # Convert to proportions
  Jones = c(92.1 / 100, 95.3 / 100, 9.4 / 100, 92.7 / 100) # Convert to proportions
)

# Melt the data for use with ggplot2
data_melted <- melt(data2, id.vars='Category')

# Create the bar plot with specified colors and corrected percentage values
ggplot(data_melted, aes(x = Category, y = value, fill = variable)) +
  geom_bar(stat = 'identity', position = position_dodge(width = 0.9)) +
  scale_fill_manual(values = c('Smith' = '#1f77b4', 'Jones' = '#d62728')) +
  theme_minimal() +
  labs(x = '', y = 'Percentage (%)', fill = '') +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = 'bottom',
    legend.title = element_blank(),
    legend.text = element_text(size = 12),
    plot.title = element_text(hjust = 0.5)
  ) +
  scale_y_continuous(labels = percent_format()) + # This will ensure y-axis is formatted as percentage
  ggtitle('Proportions for regret, social norms, and negative affect')

## Part 2: Car accident scenario

@tbl-frequency_and_proportion_exp2 shows that most participants regretted the exception (277, 81%) more strongly than the rule (65,19%; p < .001, d = 1.58). The same result can be seen more intuitively in figure2. the result for the exception event represented by Mr. White is much higher for both Regret and Luck than for the Routine event represented by Mr. Adams.

```{r}
#| echo: false
#| label: tbl-frequency_and_proportion_exp2
#| tbl-cap: "Frequency and proportion of regret and luck for exp1."
# Create the data frame based on the the value we gained from the binomial test
data <- data.frame(
  Level = c("Routine Adams", "Exception White", "Routine Adams", "Exception White"),
  Count = c(65, 277, 114, 228),
  stringsAsFactors = FALSE
)

# Split the data frame into two based on the categories: Regret and Luck
regret_data <- data[1:2, ]
luck_data <- data[3:4, ]

# Add the percentages as a new column
regret_data$Percentage <- c(19.0, 81.0)
luck_data$Percentage <- c(33.3, 66.7)

# Combine the two data frames into a single data frame for display
final_data <- data.frame(
  ` ` = c("Regret", "", "Luck", ""),
  Level = c(regret_data$Level, luck_data$Level),
  Count = c(regret_data$Count, luck_data$Count),
  Percentage = c(regret_data$Percentage, luck_data$Percentage)
)

# Create a table
kable(final_data, col.names = c("", "Choice", "Count", "%"), align = c('l', 'l', 'r', 'r'))
```

@fig-proportion_exp2 is a replication of figure2 in the original paper, showing the counts and proportions of perceived regret and luck for car accident events.

```{r}
#| echo: false
#| label: fig-proportion_exp2
#| fig-cap: "Proportion of regret and luck for exp2."

# Data in percentage for Adams and White we gained from the table we created
# Ensure that the percentages are divided by 100 to be in the correct format
data <- data.frame(
  Category = c('Regret', 'Luck'),
  Adams_Routine = c(19.0 / 100, 33.3 / 100), # Convert to proportions
  White_Exception = c(81.0 / 100, 66.7 / 100) # Convert to proportions
)

# Melt the data for use with ggplot2
data_melted <- melt(data, id.vars='Category')

# Create the bar plot with specified colors and corrected percentage values
ggplot(data_melted, aes(x = Category, y = value, fill = variable)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  scale_fill_manual(values = c('Adams_Routine' = '#1f77b4', 'White_Exception' = '#d62728')) +
  theme_minimal() +
  labs(x = '', y = 'Percentage (%)', fill = '') +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = 'bottom',
    legend.title = element_blank(),
    legend.text = element_text(size = 12),
    plot.title = element_text(hjust = 0.5)
  ) +
  scale_y_continuous(labels = percent_format()) + # This will ensure y-axis is formatted as percentage
  ggtitle('Proportions for regret and luck')

@@

@@

## Weaknesses and future directions
The main shortcomings of the study are that it does not analyze the impact of cultural differences on the perception of regret and decision-making, and the sample selected is not representative enough as it is Turkish, so the sample should be enlarged and screened to ensure that the data are of high quality.

Future research could explore how individual differences, such as personality traits, affect the experience of regret, or expand the current research methodology to comprehensively analyze the emotional responses related to regret and try to discover connections and patterns to gain a deeper understanding of this theory.

@@

## Read raw data and evaluated demographic information

We acquired a CSV file containing raw data relevant to 'Routines and Regret: An Examination of Behavioral Norms and Emotional Responses.' Subsequently, we inspected the first six rows displayed by @tbl-interested_data and assessed the demographic details.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-interested_data
#| tbl-cap: "First six rows of data of interest"

@@

#Check the top six rows of data of interest
dataset_selected <- data %>% select(Sc1_regret, sc1_socnorms1, sc1_socnorms2, sc1_combinednorms, Sc2_regret, Sc2_lucky) 
kable(head(dataset_selected))
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
# To check exclusions based on pre-registration criteria uncomment the following
# Criteria:
# Serious lower than 5
# English understanding lower than 5
# data <- data[ which(4<data$serious 
#                     & data$engunder > 4), ]

@@

## Label and test the data for statistical information

### Label the data for exp1
We label the data in the questionnaire with values of 1 and 2 as "Routine Smith", "Exception Jones" or "Routine Adams", "Exception White". The top 6 rows of interest are shown by @tbl-labeled_data_exp1 and @tbl-labeled_data_exp2.

```{r}
#| echo: false
#| warning: false
#| label: tbl-labeled_data_exp1
#| tbl-cap: "First six rows of the labeled data of interest in exp1"

@@

dataset_selected_exp1 <- data %>% select(Sc1_regret, sc1_socnorms1, sc1_socnorms2, sc1_combinednorms) 
kable(head(dataset_selected_exp1))
```

### Obtain statistical information for exp1
Statistics about the data we are interested in, such as frequencies and percentages, are generated below by the descriptives and propTest2 functions, and these statistics underpin the plotting of @tbl-frequency_and_proportion_exp1 and @fig-proportion_exp1 for experiment 1 in Results.

```{r}
#| echo: false
#| warning: false

@@

```{r}
#| echo: false
#| warning: false
#| include: false

@@

### Label the data for exp2
```{r}
#| echo: false
#| warning: false
#| label: tbl-labeled_data_exp2
#| tbl-cap: "First six rows of the labeled data of interest in exp2"
#| 
###################
# EXPERIMENT 2 (Car Accident-Scenario)

@@

dataset_selected_exp2 <- data %>% select(Sc2_regret, Sc2_lucky) 
kable(head(dataset_selected_exp2))
```

### Obtain statistical information for exp2
Statistics about the data we are interested in, such as frequencies and percentages, are generated below by the descriptives and propTest2 functions, and these statistics underpin the plotting of @tbl-frequency_and_proportion_exp2 and @fig-proportion_exp2 for experiment 2 in Results.

```{r}
#| echo: false
#| warning: false

# Descriptives for main variables
jmv::descriptives(
  data=data,
  vars=c(
    "Sc2_regret",
    "Sc2_random_1",
    "Sc2_random_2",
    "Sc2_lucky"),
  freq=TRUE)
```

```{r}
#| echo: false
#| warning: false
#| include: false
# binomial Z
jmv::propTest2(
  data=data,
  vars=c(
    "Sc2_regret",
    "Sc2_lucky"),
  ci=TRUE)
```

@@

# Questionnaire details

## Scenario #1

**Mr. Jones** almost never takes hitch-hikers in his car. Yesterday he gave a man a ride and was robbed. **Mr. Smith** frequently takes hitch-hikers in his car. Yesterday he gave a man a ride and was robbed.

### Comprehension Questions
1. Who almost never takes hitch-hikers in his car?
   - Mr. Jones (1)
   - Mr. Smith (2)
2. Who frequently takes hitch-hikers in his car?
   - Mr. Jones (1)
   - Mr. Smith (2)
3. Who got robbed?
   - Mr. Smith (1)
   - Mr. Jones (2)
   - Both Mr. Smith and Mr. Jones (4)

### Regret
- Who do you expect will experience greater regret over the episode?
  - Mr. Jones (1)
  - Mr. Smith (2)

### Social Norms 1
- Whose behavior do you think is more common in society?
  - Mr. Jones (1)
  - Mr. Smith (2)

### Social Norms 2
- Whose behavior do you think will be more criticized by others in society?
  - Mr. Jones (1)
  - Mr. Smith (2)

### Negative Affect
- Contemplating your previous answers about this scenario and factoring in both Mr. Jones and Mr. Smith's personal routines and your perceptions of social norms and possible social criticism, who do you think overall experienced more negative feelings about the decision to take a hitch-hiker that day?
  - Mr. Jones (1)
  - Mr. Smith (2)

## Scenario #2

**Mr. Adams** was involved in an accident when driving home after work on his regular route. **Mr. White** was involved in a similar accident when driving on a route that he only takes when he wants a change of scenery.

### Comprehension Questions:
1. Who was driving home after work on his regular route?
   - Mr. Adams (1)
   - Mr. White (2)
2. Who was driving on a route that he only takes when he wants a change of scenery?
   - Mr. Adams (1)
   - Mr. White (2)
3. Who was involved in an accident?
   - Mr. Adams (1)
   - Mr. White (2)
   - Both Mr. Adams and Mr. White (4)

### Regret:
- Who is more upset over the accident?
  - Mr. Adams (1)
  - Mr. White (2)

### Random Chance:
- Please rate your agreement with the following statements:
  - Mr. Adam's accident is just a random coincidence (1 – strongly disagree to 7 – strongly agree)
  - Mr. White's accident is just a random coincidence (1 – strongly disagree to 7 – strongly agree)

### Luck:
- Which of the two do you think is less lucky?
  - Mr. Adams (1)
  - Mr. White (2)
